---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

#Reproducible Research

##Course Project 1

###1. Loading and preprocessing the data

First of all, we're loading packages which we'll need for the analysis and data.

```{r results='hide'}
library(knitr)
library(dplyr)
library(lattice)

activity <- as.data.frame(read.csv("activity.csv"))
```

###2. What is mean total number of steps taken per day?

We calculate total number of steps per day.

```{r}
t01 <- activity %>% 
  group_by(date) %>%
  summarise(sumsteps = sum(steps))
```

Then, we make a histogram.

``` {r ECHO = T}
hist(t01$sumsteps, col="green", main="Steps per day", xlab="Steps per day")
```

After this, we calculate mean and median number of steps per day.

``` {r m}
mean <- mean(t01$sumsteps, na.rm=T)
median <- median(t01$sumsteps, na.rm=T)
```

So, mean is **`r mean`** 
and median is **`r median`** 

###3. What is the average daily activity pattern?

We're plotting the graph for daily acticity.

``` {r}
t02 <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(t02, type="l", xlab = "intervals", ylab = "avg. steps number")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
t02[t02==max(t02)]
```


###4. Imputing missing values

We find out what is the number of NAs.

``` {r}
sum(is.na(activity$steps))
```

Then we replace NAs with average steps number for each particular interval.

``` {r}
activity2 <- activity %>% mutate(steps = ifelse(is.na(steps), 
      t02, 
      steps))
```

After this, we're plotting for the histogram:

```{r ECHO=T}
t03 <- activity2 %>% 
  group_by(date) %>%
  summarise(sumsteps = sum(steps))

hist(t03$sumsteps, col="green", main="Steps per day", xlab="Steps per day")
```

Counting mean and median:
``` {r m2}
mean2 <- mean(t03$sumsteps)
median2 <- median(t03$sumsteps)
```

So, mean is **`r mean2`** 
and median is **`r median2`** 

Imputting missing data made median equal to mean.

###5. Are there differences in activity patterns between weekdays and weekends?

We create variable for weekdays:

``` {r}
activity2$date <- as.Date(as.character(activity2$date))

activity3 <- activity2 %>% mutate(date = as.Date(as.character(date))) %>%
                                 mutate(weekday = as.factor(ifelse(
                                   as.numeric(as.factor(weekdays(activity2$date))) %in% c(6,7),
                                      "weekend",
                                      "weekday")))
```

and plot for the differences between weekdays and weekends.

```{r}
activity4<- activity3 %>% 
     group_by(weekday,interval) %>%
     summarise(meansteps=mean(steps))
with (activity4, 
             xyplot(meansteps ~ interval|weekday, type="l", 
                                   ylab="Number of steps",layout=c(1,2)))      
```

Daily activity on weekdays and weekends has differences.
